# Data

## 2.1 Technical description
The World Income Inequality Database (WIID) contains comprehensive information about income inequality for 201 countries and historical entities. The current version contains over 24,000 data points with observations from pre-1960s to 2022. The database combines data from various authoritative sources including:

1. LIS Cross-National Data Center (Luxembourg Income Study)
2. Eurostat
3. Socio-Economic Database for Latin America and the Caribbean (SEDLAC)
4. United Nations
5. National statistical offices
6. OECD
7. World Bank's Poverty and Inequality Platform (PIP)
8. Research studies
9. Other international organizations

The observations are distributed across time periods as follows:

| Time span | Number of observations |
|:----------|---------------------:|
| Total observations | 24,367 |
| Before 1960 | 311 |
| 1960-69 | 714 |
| 1970-79 | 946 |
| 1980-89 | 1,651 |
| 1990-99 | 3,758 |
| 2000-09 | 6,764 |
| 2010-19 | 8,753 |
| 2020- | 1,470 |




## 2.2 Missing value analysis

We want to first look at the range of years for the data by country and by continent to see if we can observe any patterns.
```{r}
library(readxl)
library(ggplot2)
library(tidyr)
data <- read_excel("WIID_28NOV2023.xlsx")
data <- data %>%
  mutate(region_un = ifelse(country == "Turkiye", "Europe", region_un))

```

```{r, fig.height=8, fig.width=18}
data <- data %>%
  mutate(country = factor(country, levels = data %>%
                            distinct(country, region_un) %>%
                            arrange(region_un, country) %>%
                            pull(country)))

# Create the plot without connecting points and retaining multiple data points per country
ggplot(data, aes(x = country, y = year, color = region_un)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Year Range of Data for Each Country", 
       x = "Country", 
       y = "Year", 
       color = "Region")

```
From the plot, we can observe that even though the data spans over 100 years, most records are concentrated between the 1950s and 2020s. We can also see that countries in Europe and America have the most frequent data recordings, with almost a data point for each year. Whereas for African countries, data are measured less frequently.